---
title: "Pumpkins"
author: "Edward Hill"
date: "11-12-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

Let's train a model for[giant pumpkins](rfordatascience/tidytuesday), predicting a giant pumpkin weight from other characteristics

## Explore Data

```{r}
library(tidyverse)

pumpkins_raw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-19/pumpkins.csv")

pumpkins <-
  pumpkins_raw %>%
  separate(id, into = c("year", "type")) %>%
  mutate(across(c(year, weight_lbs, ott, place), parse_number)) %>%
  filter(type == "P") %>%
  select(weight_lbs, year, place, ott, gpc_site, country)
pumpkins
```

The main relationship here is between the volume/size of the pumpkin (measured via “over-the-top inches”) and weight.

```{r}
pumpkins %>%
  filter(ott > 20, ott < 1e3) %>%
  ggplot(aes(ott, weight_lbs, color = place)) +
  geom_point(alpha = 0.2, size = 1.1) +
  labs(x = "over-the-top inches", y = "weight (lbs)") +
  scale_color_viridis_c()
```

Big, heavy pumpkins placed closer to winning at the competitions, naturally!

Has there been any shift in this relationship over time?

```{r}
pumpkins %>%
  filter(ott > 20, ott < 1e3) %>%
  ggplot(aes(ott, weight_lbs)) +
  geom_point(alpha = 0.2, size = 1.1, color = "gray60") +
  geom_smooth(aes(color = factor(year)),
    method = lm, formula = y ~ splines::bs(x, 3),
    se = FALSE, size = 1.5, alpha = 0.6
  ) +
  labs(x = "over-the-top inches", y = "weight (lbs)", color = NULL) +
  scale_color_viridis_d()
```

Hard to say, I think.

Which countries produced more or less massive pumpkins?

```{r}
pumpkins %>%
  mutate(
    country = fct_lump(country, n = 10),
    country = fct_reorder(country, weight_lbs)
  ) %>%
  ggplot(aes(country, weight_lbs, color = country)) +
  geom_boxplot(outlier.colour = NA) +
  geom_jitter(alpha = 0.1, width = 0.15) +
  labs(x = NULL, y = "weight (lbs)") +
  theme(legend.position = "none")
```

## Build and fit a workflow set

```{r}
library(tidymodels)
set.seed(123)
pumpkin_split <- pumpkins %>%
  filter(ott > 20, ott < 1e3) %>%
  initial_split(strata = weight_lbs)

pumpkin_train <- training(pumpkin_split)
pumpkin_test <- testing(pumpkin_split)

set.seed(234)
pumpkin_folds <- vfold_cv(pumpkin_train, strata = weight_lbs)
pumpkin_folds
```

Next, let’s create three data preprocessing recipes: one that only pools infrequently used factors levels, one that also creates indicator variables, and finally one that also creates spline terms for over-the-top inches.

```{r}
base_rec <-
  recipe(weight_lbs ~ ott + year + country + gpc_site,
    data = pumpkin_train) %>%
  step_other(country, gpc_site, threshold = 0.02)

ind_rec <-
  base_rec %>%
  step_dummy(all_nominal_predictors())

spline_rec <-
  ind_rec %>%
  step_bs(ott)
```

Then, let’s create three model specifications: a random forest model, a MARS model, and a linear model.

```{r}
rf_spec <-
  rand_forest(trees = 1e3) %>%
  set_mode("regression") %>%
  set_engine("ranger")

mars_spec <-
  mars() %>%
  set_mode("regression") %>%
  set_engine("earth")

lm_spec <- linear_reg()
```

Now it’s time to put the preprocessing and models together in a workflow_set().

```{r}
pumpkin_set <-
  workflow_set(
    list(base_rec, ind_rec, spline_rec),
    list(rf_spec, mars_spec, lm_spec),
    cross = FALSE)
pumpkin_set

```

We use cross = FALSE because we don’t want every combination of these components, only three options to try. Let’s fit these possible candidates to our resamples to see which one performs best.

```{r}
library(doParallel)
library(ranger)
library(earth)
registerDoParallel()
set.seed(2021)

pumpkin_rs <-
  workflow_map(
    pumpkin_set,
    "fit_resamples",
    resamples = pumpkin_folds
  )

pumpkin_rs
```


## Evaluate workflow set

How did our three candidates do?

```{r}
autoplot(pumpkin_rs)
```

There is not much difference between the three options, and if anything, our linear model with spline feature engineering maybe did better. This is nice because it’s a simpler model!

```{r}
collect_metrics(pumpkin_rs)
```

We can extract the workflow we want to use and fit it to our training data.

```{r}
final_fit <-
  extract_workflow(pumpkin_rs, "recipe_3_linear_reg") %>%
  fit(pumpkin_train)
tidy(final_fit)
```

We can use an object like this to predict, such as on the test data like predict(final_fit, pumpkin_test), or we can examine the model parameters.

```{r}
tidy(final_fit) %>%
  arrange(-abs(estimate))
```

The spline terms are by far the most important, but we do see evidence of certain sites and countries being predictive of weight (either up or down) as well as a small trend of heavier pumpkins with year.

Don’t forget to take the tidymodels survey for 2022 priorities!





