---
title: "Age of the Universe using linear regression"
author: "Edward Hill"
date: "19-12-2021"
output: 
  html_document:
      theme: journal
bibliography: R_GAM.bib
biblio-style: apalike
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The big-bang model implies that the universe expands uniformly according to Hubble's law:[^note1] 

<math display = 'block'>
  <mrow>
    <mi>y</mi>
    <mtext fontfamily='Times New Roman'>&nbsp;</mtext>
    <mo>=</mo>
    <mtext fontfamily='Times New Roman'>&nbsp;</mtext>
    <mi>&beta;</mi>
    <mi>x</mi>
  </mrow>
</math>

Where <mi>y</mi> is the relative velocity of any two galaxies seperated by distance <mi>x</mi>, and <mi>&beta;</mi> is "Hubble's constant". 

<mi>&beta;</mi> is the approximate age of the universe, but <mi>&beta;</mi> is unknown and must be estimated from observations of <mi>x</mi> and <mi>y</mi>. 

See; @Wood2017, p. 1-9

## Data hubble: 

```{r, echo=FALSE}
library(gamair) # contains 'hubble'
data(hubble)
hubble
hub.mod <- lm(y~x-1,data=hubble)
summary(hub.mod)
```

In this line I'll give you the mean-value of y: `r round(mean(hubble$y),1)`. 

### Plot the residuals against the fitted values

```{r}
plot(fitted(hub.mod),residuals(hub.mod),xlab="fitted values",ylab="residuals")
```

### Omit offending points and produce new residual plot

```{r}
hub.mod1 <- lm(y~x-1,data=hubble[-c(3,15),])
summary(hub.mod1)
plot(fitted(hub.mod1),residuals(hub.mod1),xlab="fitted values",ylab="residuals")
```

### Estimate Hubble's Constant

```{r}
hubble.const <- c(coef(hub.mod),coef(hub.mod1))/3.09e19
age <- 1/hubble.const
age
age/(60^2*24*365)
```

### 1.1.3 Adding a distributional assumption

### Testing Hypotheses about <mi>&beta;</mi> in Creationist Theory

```{r}
cs.hubble <- 163000000
t.stat<-(coef(hub.mod1)-cs.hubble)/summary(hub.mod1)$coefficients[2]
pt(t.stat,df=21)*2
```

### Confidence intervals

```{r}
sigb <- summary(hub.mod1)$coefficients[2]
h.ci<-coef(hub.mod1)+qt(c(0.025,0.975),df=21)*sigb
h.ci
h.ci<-h.ci*60^2*24*365.25/3.09e19 # convert to 1/years
sort(1/h.ci)
```

[^note1]: Easiest way of writing mathematical equation in R Markdown  See Youtube: https://www.youtube.com/watch?v=4I3PCDME5U8)

Relevant literature;

1. [@Wood2017]

2. [@Crawley2013]

3. [How much more likely? The Implications for Odds Ratios for Probabilities by Liberman](https://github.com/EdwardHill15/Udemy_GAM/blob/main/Linear%20Regression%20GLMs%20and%20GAMs%20with%20R/other%20documentation/liberman_2005_OR_paper_AJE.PDF)
[@Liberman2005]

# References