---
title: "Age of the Universe using linear regression"
author: "Edward Hill"
date: "19-12-2021"
output: 
  html_document:
      theme: journal
bibliography: R_GAM.bib
biblio-style: apalike
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The big-bang model implies that the universe expands uniformly according to [Hubble's law](https://en.wikipedia.org/wiki/Hubble%27s_law).

<math display = 'block'>
  <mrow>
    <mi>y</mi>
    <mtext fontfamily='Times New Roman'>&nbsp;</mtext>
    <mo>=</mo>
    <mtext fontfamily='Times New Roman'>&nbsp;</mtext>
    <mi>&beta;</mi>
    <mi>x</mi>
  </mrow>
</math>

Where <mi>y</mi> is the relative velocity of any two galaxies seperated by distance <mi>x</mi>, and <mi>&beta;</mi> is "Hubble's constant". in standard astrophysical notation <mi>y</mi> = <mi>v</mi>, <mi>x</mi> = <mi>d</mi> and <mi>&beta;</mi> = <mi>H0</mi>.

<mi>&beta;</mi> is the approximate age of the universe, but <mi>&beta;</mi> is unknown and must be estimated from observations of <mi>x</mi> and <mi>y</mi>. 

See; @Wood2017, p. 1-9

## Data hubble: 

```{r, echo=FALSE}
library(gamair) # contains 'hubble'
data(hubble)
hubble
hub.mod <- lm(y~x-1,data=hubble)
summary(hub.mod)
```

[Hubble's Law](https://en.wikipedia.org/wiki/Hubble%27s_law), also known as the [Hubble-Lemaitre law](https://www.iau.org/news/pressreleases/detail/iau1812/?lang), is the observation in physical cosmology that galaxies are moving away from Earth at speeds proportional to their distance. In other words, the farther they are the faster they are moving away from Earth. The velocity of the galaxies has been determined by their redshift, a shift of the light they emit toward the red end of the visible spectrum.

Hubble constant is most frequently quoted in (km/s)/Mpc, thus giving the speed in km/s of a galaxy 1 megaparsec (3.09 time 1019 km) away, and its value is about 70 (km/s)/Mpc. However, the SI unit of H0 is simply s^-1^, and the SI unit for the reciprocal of H0 is simply the second. The reciprocal of H0 is known as the Hubble time. The Hubble constant can also be interpreted as the relative rate of expansion. In this form H0 = 7%/Gyr, meaning that at the current rate of expansion it takes a billion years for an unbound structure to grow by 7%.

### Plot the residuals against the fitted values

```{r}
plot(fitted(hub.mod),residuals(hub.mod),xlab="fitted values",ylab="residuals")
```

### Omit offending points and produce new residual plot

```{r}
hub.mod1 <- lm(y~x-1,data=hubble[-c(3,15),])
summary(hub.mod1)
plot(fitted(hub.mod1),residuals(hub.mod1),xlab="fitted values",ylab="residuals")
```

### Estimate Hubble's Constant

```{r}
hubble.const <- c(coef(hub.mod),coef(hub.mod1))/3.09e19
age <- 1/hubble.const
age
age/(60^2*24*365)
```


### Testing Hypotheses about <mi>&beta;</mi> in Creationist Theory

```{r}
cs.hubble <- 163000000
t.stat<-(coef(hub.mod1)-cs.hubble)/summary(hub.mod1)$coefficients[2]
pt(t.stat,df=21)*2
```

### Confidence intervals

```{r}
sigb <- summary(hub.mod1)$coefficients[2]
h.ci<-coef(hub.mod1)+qt(c(0.025,0.975),df=21)*sigb
h.ci
h.ci<-h.ci*60^2*24*365.25/3.09e19 # convert to 1/years
sort(1/h.ci)
```

Mean-value for <mi>v</mi> or <mi>y</mi> = `r round(mean(hubble$y),0)` km/s. 

[^note1]: Easiest way of writing mathematical equation in R Markdown  See Youtube: https://www.youtube.com/watch?v=4I3PCDME5U8)

Relevant literature;

1. [@Wood2017]

2. [@Crawley2013]

3. [How much more likely? The Implications for Odds Ratios for Probabilities by Liberman](https://github.com/EdwardHill15/Udemy_GAM/blob/main/Linear%20Regression%20GLMs%20and%20GAMs%20with%20R/other%20documentation/liberman_2005_OR_paper_AJE.PDF)
[@Liberman2005]

# References